---
title: "Fastp"
author: "Thomas Källman"
date: "7/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fastp

The short read filter tool fastp can can be used for extracting UMIs (unique molecular identifiers) that are part of many single cell or RNA-seq data sets and allow for removing PCR artifacts etc as all sequences will start with a short unique signature tag.

Some sequence libraries created for eDNA can use this functionality to effeciently parse tags from the start of reads and hence extract tags from the beginning of reads and add the tag to the name of the reads.

```{bash}
fastp -i P8111_1001_S1_L001_R1_001.fastq.gz -I P8111_1001_S1_L001_R2_001.fastq.gz -o P8111_1001_S1_L001_R1_001_fastp1.fastq.gz -O P8111_1001_S1_L001_R2_001_fastp1.fastq.gz -A -Q -L -U --umi_loc=read1 --umi_len=8

fastp -i P8111_1001_S1_L001_R1_001_fastp1.fastq.gz -I P8111_1001_S1_L001_R2_001_fastp1.fastq.gz -o P8111_1001_S1_L001_R1_001_fastp2.fastq.gz -O P8111_1001_S1_L001_R2_001_fastp2.fastq.gz -A -Q -L -U --umi_loc=read2 --umi_len=8

```


This will move the start of both the forward and reverse reads to the name of the reads. The read names will after this contain:

tagForward:tagReverse in the names of both the forward and reverse reads

## Cutadapt
Using cutadapt reads are retained only if the contain the forward and reverse amplification primer. We also trim these from the reads hence leaving only actual sequence data of interest for each read.

```{bash}
cutadapt -a RGACGAGAAGACCCTATARA...TACYTTAGGGATAACAGCGT -A ACGCTGTTATCCCTAARGTA...TYTATAGGGTCTTCTCGTCY --discard-untrimmed -o out.1.fastq.gz -p out.2.fastq.gz P8111_1001/02-FASTQ/170519_M01548_0122_000000000-B5F5V/P8111_1001_S1_L001_R1_001_test2.fastq.gz P8111_1001/02-FASTQ/170519_M01548_0122_000000000-B5F5V/P8111_1001_S1_L001_R2_001_test2.fastq.gz -m 50
```

We hence now have new fastq files with forward and reverse reads that should only harbor the sequence of interest and have tags as part of their read names.
We can import this into R using the ShortRead package and then look at the tags obtained in the sequence runs.

```{r}
library(ShortRead)
```

```{r}
sampler <- FastqSampler('/Users/thokall/Soil/out.1.2.fastq.gz',
			  100000)
  set.seed(13) # for consistency in evaluation of procedures
  read1 <- yield(sampler)

```


```{r}
trimFilter <- function(fq1, fq2, tags, destination = "out") {
  ## open input stream
  #stream <- open(FastqStreamer(fl))
  stream_R1 <- ShortRead::FastqStreamer(fq1)
  stream_R2 <- ShortRead::FastqStreamer(fq2)
  on.exit(close(stream_R1))
  on.exit(close(stream_R2), add = TRUE)
  #on.exit(close(stream))
  repeat {
    ## input chunk
    fq_R1 <- yield(stream_R1)
    fq_R2 <- yield(stream_R2)
    if (length(fq_R1) == 0)
      break
    sel <- grepl(tags, ShortRead::id(fq_R1))
    writeFastq(fq_R1[sel], file = gsub("../Soil", replacement = "../Soil/out", x = fq1, fixed = TRUE), mode = "a")
    writeFastq(fq_R2[sel], file = gsub("../Soil", replacement = "../Soil/out", x = fq2, fixed = TRUE), mode = "a")
    
  }
}
```